### 2015-10-06 src/table-synthetic.tex
experiment: add table-synthetic.tex.

\documentclass{standalone}
\usepackage{graphicx}
\begin{document}

    \begin{tabular}{|c|c|c|}
        {\bf 2 hyperplanes} &
        {\bf 3 hyperplanes} &
        {\bf 4 hyperplanes}
        \\
        \multicolumn{3}{|c|}{{\bf Deterministic, single margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Deterministic, multiple margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, single margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, multiple margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, single margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, multiple margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png}
    \end{tabular}
\end{document}