### 2016-03-11 paper/rebuttal.txt
paolo: commenti a reviewer 1

We would like to thank all reviewers for their feedback. We will buy an
additional page for the camera ready to enrich the related works section and
improve the text where it was previously cut due to space limitations.
Additional replies:

REVIEW 1

% - novelty
%     - certamente le alternative/sets non sono nuovi; la novita' e' la
%       suddivisione dello spazio di ricerca
%     - cercare roba recente che cita Guo/Bonilla/Viappiani e dimostrare che
%       non esiste nulla di piu' sul punto
%     - diversity in IR esiste [ma non ha lo stesso scopo (non dipende da w)] ma
%       per spazio ci siamo concentrati su related work piu' vicini; si potrebbe
%       comprare una pagina ulteriore nel camera ready per spiegare.
% 
% - set search not analyzed thoroughly
%     - il metodo e' giustificato anche con k=2, che funziona pure bene
%     - k>2 e' utile in base al costo cognitivo; per questo abbiamo messo 2
%       grafici (per query e per iteration); qui si vede il vantaggio di k>2
%     - no messaggio definitivo per via del costo cognitivo e del modello utente;
%       si possono esplorare alternative ma non e' il nostro fuoco; o vederlo
%       come recommendation set (di 5 prendi 1), pero' avvantaggerebbe il nostro
%       metodo
% 
% - *
%     - il fatto che i Bayes siano piu' lenti sono comunque lo stato dell'arte:
%       c'e' un tradeoff tra numero di domande e scalare nel bayesiano; noi
%       invece otteniamo le stesse perf ma scaliamo molto oltre.
%     - gli attributi booleani c'e' un misunderstanding: il problema e' ben
%       piu' complesso. spiegare attributi in dettaglio e quanti bits.

- First of all, there is a misunderstanding regarding the experiment with the PC dataset. 
The choice space is much larger than perceived by the reviewer: the binary encoding of the 8
features (7 discrete + 1 real) uses 73 boolean attributes. This corresponds to more than 700K feasible configurations.

- While the idea of using sets of candidates to encourage diversity is not new, the
  novelty is in extending the large margin paradigm to sets. We view this as a significant step forward over previous approaches using max-margin maximization and heuristic methods for uncertainty reduction [for example: Gajos 2005].

- Diversity in information retrieval has a different goal, unrelated to
  optimization and learning. We will provide details and further references
  in the additional page. QUESTO LO LEVEREI

- Recently the focus in preference elicitation shifted a bit towards sequential decision making and preference-based reinforcement learning. Our literature search found nothing more appropriate than the works we compare to. We'd be grateful for pointers to other relevant references.

- The main contribution of the work is that splitting the search space
  in the fashion proposed achieves state-of-the-art performance at a
  fraction of the computational cost. This is certainly true for
  pairwise queries (k=2), where all methods can be compared on a fair
  basis. For k>2, it is difficult to precisely define the cognitive
  load.  That's why we report both per-query and per-iteration 
  results, where the latter show faster convergence for k>2.

- The fact that Bayesian methods are computationally expensive is not
  sufficient to rule them out, as they tend to be more accurate on
  small scale problems. What we show is that we are able to get
  comparable performance on these problems, while being able to scale much more.

REVIEW 2

% R2
% 
% - citare UTA in OR -> citare UTA quando introduciamo le slacks
% - ha ragione, rephrase (primo qualita', secondo diversita', tutti e due diversita' dei buoni)
% - non normalizzando i pesi (che e' solo una possibilita'), la L1 e' lo standard
%   per sparsificare
% - explain that the linearization trick is used in Boutilier et al. 2006
% - chiarire che il num bool e sum domini; ma i problemi sono tranquilli perche'
%   i w utenti sono semplici e gli XOR restringono il problema
% - why should gamma be crossvalidated? intuition:
%     - il vincolo 2 diversifica
%     - il gamma vuole che gli oggetti siano di buona qualita' che in PE servono
%       perche' l'utente non si incazzi; non e' ovvio quale deve essere il suo
%       contributo percio' crossvalidiamo
%     - chiarire i bump a cosa sono dovuti (non a gamma); tagliato per spazio
% - fix footnote 4 only applies to k=1 (chiarisco che chiariro')
% - explain why 80% sparsity mimics a user (typically cares about 3 or 4 features)
%     - siamo d'accordo che con 3-4 features non serve la sparsita'; peccato
%       che dobbiamo simularlo su poche features (synth) for sake of comparisons
%       perche su PC gli altri metodi non funzionano
%     - la sparsita' non l'abbiamo introdotta per far vedere che il nostro sys
%       va meglio -- l'idea e' che scaliamo meglio sia in sparso che in denso.
%       ma usarla e' ragionevole per modellare l'utente. non ci costa nulla
%       toglierla, e non cambia la sua utilita'. il confronto non e' sulla
%       sparsita' ma sul fatto che gli altri non scalano)
%       (infatti la scalabilita' e' il nostro cavallo di battaglia)

- Thanks for the reference, it will be integrated.
- Correct, <wi,xi> encourages the quality of the xi's, diversity is given
  by constraint 2. We will rephrase the sentence.
- Fixing the 1-norm is not the only option. In our case, the norm is variable
  and sparsity is encouraged through LASSO. This is a standard approach.
- We will clarify that the linearization trick is taken from elsewhere, and
  refer to the paper by Boutilier.
- We plan to extend the work to real variables, as per the Conclusion;
  but this is beyond the scope of this paper.
- User resources strongly limit the overall complexity of the MILP problem.
  Current MILP solvers can easily deal with problems of this magnitude.
- Gamma controls the importance of candidate item quality for ensuring
  diversity; gamma=0 could be used. Since the correlation is not known
  beforehand, we include gamma in the cross-validation.
- Footnote 4 should only apply to k=1 (as constraint 2 vanishes. Will fix.
- NOT SURE HOW TO WRITE IT DOWN

REVIEW 3

% - citeremo e commenteremo, grazie
% - margin must be non-negative, make clearer

- Thanks for the references, we will make sure to include them.
- The margin must be non-negative, as specified in the text above Eq 1; we will
  point it out more clearly.

REVIEW 4

% - BT: forse ci favorisce, ci stiamo solo confrontando con quello che fanno gli
%   altri, preso dai competitors; al contrario, sono i modelli bayesiani che
%   lo sfruttano meglio (o almeno che potrebbero)
% - cognitive load: ha ragione, ma tutti fanno domande a coppie es Guo & Sanner,
%   ci accodiamo. non e' chiaro come formalizzare (chiedere agli psicologi)? e'
%   da investigare. in effetti e' quello che vogliamo fare (trovare query che
%   sono piu' facili per l'utente). ma va oltre lo scopo.
% - scalability: stiamo sempre parlando di interazione con l'utente, quindi siamo
%   lontani dal raggiugnere la complex gestibile da un MILP solver. il PC ad es
%   e' gestibile mentre per l'utente e' complesso (non si va sulle migliaia di
%   var booleane). capire quanti Bools un MILP gestisce (chiedere a Roberto?).
%   la sat del solver sta molto sopra la sat dell'utente.
% - reals: future work di sicuro, infatti l'abbiamo messo li' -- le estensioni
%   sarebbero non-lineari per via di <w,x>, citare ultimo paper sanner. in
%   realta' ongoing su hybrid (alternated o nonlinear).

%- We do not think the BT response model favors our method. First, it was
%  previously used to evaluate the competitors, we merely took it from there.
%  Second, given that Bayesian models are tied to the response model, they are
%  more likely to benefit from the additional constraints that the modelling
%  provides.

- We adopt Bradley Terry (BT) as a response model as it is a common assumption in preference elicitation papers. It is true that constraint 2 induce items that are "different" and will often be associated with a low prob of error according to the BT reponse model.
But we do not think this as giving an unfair advantage to our method; actually it is the Bayesian competitors that exploit BT better, as it is used as an observation model when performing Bayesian updates.

- True, modelling the cognitive load is not easy. We have ongoing work to
  explore this issue further, in collaboration with psychologists. In the
  paper we just followed the common practice of pairwise queries, for
  comparability.
- User resources strongly limit the overall complexity of the MILP problem;
  MILP solvers are well-equipped to solve problems of this magnitude.
- Indeed linearly dependent variables are not very expressive. We are very much
  interested in extending the method to hybrid Boolean/continuous domains, as
  hinted to in the Conclusion. We are currently investigating two directions:
  alternating between optimization of Boolean and continuous variables, and
  leveraging non-linear solvers.
- Thanks for the reference, we'll make sure to include it.