### 2016-03-11 paper/rebuttal.txt
rebuttal: removed -

We would like to thank all reviewers for their feedback. Detailed replies follow:

REVIEW_1

First of all, there is a misunderstanding regarding the experiment
with the PC dataset.  The choice space is much larger than perceived
by the reviewer: the binary encoding of the 8 features (7 discrete, 1
real) uses 73 boolean attributes. This corresponds to more than 700K
feasible configurations.

While the idea of using sets of candidates to encourage diversity is
not new, the novelty is in extending the large margin paradigm to
sets. We view this as a significant step forward over previous
approaches using maxmargin maximization and heuristic methods for
uncertainty reduction.

Recently the focus in preference elicitation shifted a bit towards
sequential decision making and reinforcement learning. Our literature
search found nothing more appropriate than the works we compare
to. We'd be grateful for pointers to other relevant references.

The main contribution of the work is that our method for splitting the
search space achieves state-of-the-art performance at a fraction of
the computational cost. This is certainly true for pairwise queries
(k=2), where all methods can be compared.  For k>2, it is difficult to
precisely define the cognitive load.  That's why we report both
perQuery and perIiteration results; the latter show faster
convergence for k>2.

The fact that Bayesian methods are computationally expensive is not
sufficient to rule them out, as they tend to be more accurate on small
problems. Our method achieves comparable performance on these
problems, while capable of scaling much more.

REVIEW_2

Thanks for the reference, it will be integrated.

Correct, <wi,xi> encourages the quality of the xi's; diversity is
given by constraint 2. We will rephrase.

Fixing the 1_norm is not the only option. In our case, the norm is
varying and sparsity is encouraged through LASSO. This is a standard
approach.

We will clarify that the linearization trick is a classic method in OR.

Our plan for extending the approach to real variables is not to binarize
them, but rather use appropriate solvers like MIQP.

Gamma controls the importance given to the quality of the candidate
(we want to suggest 'good' items). This needs to be traded with respect
to diversity, that's why we include gamma in the crossvalidation.

Footnote 4 is only relevant to the case k=1 (as constraint 2 vanishes). Will
fix.

The intuition for the sparse utility is based on human's bounded
rationality in large decision spaces. We agree 3 or 4
features do not require sparsity, but these were the domain sizes
suitable for the comparison with Bayesian methods. Please note that
the scalability of our method does not depend on the sparsity
assumption.

REVIEW_3

Thanks for the references, we'll include them.

The margin must be nonnegative, as specified in the text above Eq.1;
we will point it out more clearly.

REVIEW_4

Bradley Terry (BT) is a common assumption in preference
elicitation. It is true that constraint 2 tends to induce items with a
low error probability according to BT. However we do not see this as
giving us an unfair advantage; actually it is the Bayesian competitors
that exploit BT better when using it for performing Bayesian updates.

Modelling the cognitive load is indeed not easy, we just followed the
common practice of pairwise queries. We have ongoing work in
collaboration with psychologists to explore this issue further.

User cognitive resources (rather than computational ones) limit the
size of problems; MILP solvers can easily address optimizations of this
magnitude.

Indeed linearly dependent variables are not very expressive. We are
interested in extending to hybrid Boolean/continuous domains.

We'll include the suggested reference.